---
title: "Multilevel Bayesian analysis with multiple imputation"
author: "Butovens Médé"
date: "3/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("tidyverse", "skimr", "mice", "brms", "dplyr", "VIM", "naniar", "performance", "see", "qqplotr", "randomforest", "bayesplot")
library(tidyverse)
library(skimr)
library(mice)
library(brms)
library(dplyr)
library(VIM)
library(naniar)
library(performance)
library(see)
library(qqplotr)
library(randomForest)
library(bayesplot)
```


### Import data
```{r}
# Import data
SSAP <- read_csv(file.choose()) # choose SSAP_melody_data.csv file

# Look at data
skim(SSAP)

# change names
SSAP <- rename(SSAP, ID = `Imitation 3`)

# Look at missing data pattern
# md.pattern(SSAP)
SSAP %>% aggr()
SSAP %>% filter(ID != "MBUCM_13") %>% aggr()
```


### Look at missing data pattern
```{r}
# Look at missing data pattern
S_pivot_filt %>% naniar::vis_miss()
SSAP %>% filter(ID != "MBUCM_13") %>% naniar::vis_miss()

aggr_plot <- aggr(SSAP_pivot, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(SSAP_pivot), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

### Add covariates to data set
```{r}
### Add gender column to data set
SSAP <- SSAP %>% 
  mutate(Gender = case_when(str_detect(ID, pattern = "_T1") ~ "M",
                            str_detect(ID, pattern = "_10") ~ "M",
                            str_detect(ID, pattern = "_12") ~ "M",
                            TRUE ~ "F"))


### Add age for each participant
SSAP <- SSAP %>% 
  mutate(Age = case_when(str_detect(ID, pattern = "T1") ~ 34,
                         str_detect(ID, pattern = "T2") ~ 23,
                         str_detect(ID, pattern = "02") ~ 18,
                         str_detect(ID, pattern = "03") ~ 19,
                         str_detect(ID, pattern = "04") ~ 20,
                         str_detect(ID, pattern = "10") ~ 18,
                         str_detect(ID, pattern = "11") ~ 20,
                         str_detect(ID, pattern = "12") ~ 23,
                         str_detect(ID, pattern = "13") ~ 19,
                         TRUE ~ 22))


### Add last practiced age for each participant
SSAP <- SSAP %>% 
  mutate(Last_practiced = case_when(str_detect(ID, pattern = "T1") ~ 13,
                         str_detect(ID, pattern = "T2") ~ 20,
                         str_detect(ID, pattern = "02") ~ 14,
                         str_detect(ID, pattern = "03") ~ 17,
                         str_detect(ID, pattern = "04") ~ 12,
                         str_detect(ID, pattern = "07") ~ 19,
                         str_detect(ID, pattern = "10") ~ 18,
                         str_detect(ID, pattern = "11") ~ 12,
                         str_detect(ID, pattern = "12") ~ 8,
                         str_detect(ID, pattern = "13") ~ 12, 
                         TRUE ~ 17))

#### Add number of years since last practice
SSAP <- SSAP %>% 
  mutate(Since_last = Age - Last_practiced)


### Test for MCAR: little's test
SSAP  %>% select(!Last_practiced) %>%  mcar_test() # Remove Last practiced column from analysis because it is perfectly correlated with Age and Since Last, thus causing error in the evaluation of the MCAR

### Change ID, Gender to factor
SSAP <- SSAP %>% 
  mutate(ID = factor(ID),
        # Gender = factor(Gender),
         Interval = factor(Interval))

### Contrasts Gender
SSAP$Gender %>% contrasts()

### Create missingness indicator for all the variables with missing data
r_SSAP <- SSAP %>% mutate(r_prS = if_else(is.na(PRE_S),1,0 ),
                        r_poS = if_else(is.na(POST_S),1,0 ),
                        r_prV = if_else(is.na(PRE_V),1,0 ),
                        r_poV = if_else(is.na(POST_V),1,0 ),
                        r_prLC = if_else(is.na(PRE_LC),1,0 ),
                        r_poLC = if_else(is.na(POST_LC),1,0 ),
                        r_prRC = if_else(is.na(PRE_RC),1,0 ),
                        r_poRC = if_else(is.na(POST_RC),1,0 ))
```


### Change data layout for analysis
```{r}

### Pivot column
SSAP_pivot <- SSAP %>% 
  pivot_longer(cols = PRE_S:POST_LC,
               names_to = "Condition",
               values_to = "Values") %>% 
  mutate(Post = if_else(str_detect(Condition, pattern = "POST"), "POST", "PRE"),
         Condition = case_when(str_detect(Condition, pattern = "_S") ~ "Sham",
                               str_detect(Condition, pattern = "_V") ~ "Visual",
                               str_detect(Condition, pattern = "_RC") ~ "RCereblm",
                               TRUE ~ "LCereblm"))
skim(SSAP_pivot)
### Transform Condition as factor
SSAP_pivot <- SSAP_pivot %>% 
  mutate(Condition = factor(Condition),
         Post = factor(Post))

### Check contrasts for Condition (remember to change to the desired contrast)
SSAP_pivot$Condition %>% contrasts()
SSAP_pivot$Post %>% contrasts()
SSAP_pivot$Gender %>% contrasts()

### Change contrasts coding
contrasts(SSAP_pivot$Condition) <- cbind(c(1, 0, 0, 0),
                                        c(0, 1, 0, 0),
                                        c(0, 0, 0, 1))
contrasts(SSAP_pivot$Gender) <- c(-0.5, 0.5)
contrasts(SSAP_pivot$Post) <- c(1, 0)

### Mean center continuous variables
SSAP_pivot <- SSAP_pivot %>% 
  mutate(Age_c = scale(x = Age, center = T, scale = F),
          Last_practiced_c = scale(x = Last_practiced, center = T, scale = F),
          Values_c = scale(x = Values, center = T, scale = F))




### MIssing data pattern after data wrangling
SSAP_pivot %>% aggr()
SSAP_pivot %>% naniar::vis_miss()
ggsave("missing.png")
SSAP_pivot %>% prop_miss()
```


### Try different models before imputation
```{r}

### Check normality assumption
mod <- brm(Values ~ Condition * Post + Age  + Since_last + Gender + (1 + Condition + Post | ID),
           sample_prior = TRUE,
           data = SSAP_pivot)

mod.1 <- brm(Values ~ Condition * Post + Age  + Since_last + (1 + Condition | ID),
                  data = SSAP_pivot)

mod.2 <- brm(Values ~ Condition * Post + Age  + Since_last + (1 + Post | ID),
                  data = SSAP_pivot)

mod1 <- lme4::lmer(Values ~ Condition + Post + (1 + Condition | ID),
                  data = S_pivot_filt)

mod2 <- lme4::lmer(Values ~ Condition * Post + (1 + Condition + Post | ID),
                  data = S_pivot_filt)

lm1 <- lm(Values ~ Condition + Post + Age  + Since_last + Gender,
                  data = S_pivot_filt)
### Create data set wihout participants 13 who have most of its missing values
S_pivot_filt <- SSAP_pivot %>% 
  filter(ID != "BMUCM_13")

### Check some of the models created
summary(mod1)
model_performance(mod1)
check_convergence(mod.2)
check_collinearity(mod.2)
check_normality(mod.2)
check_outliers(mod.2)


### Look at potential distribution to be used with the model
check_distribution(mod.2)

### Compare models
comp <- compare_performance(lm, lm1)
comp
```


### Plot historgram for deviation scores
```{r}

### plot histogram of Values for all 4 conditions Pre and Post mixed together
ggplot(SSAP_pivot %>% filter(Condition == "LCereblm")) +
  geom_histogram( aes(x = Values), bins = 100) +
  geom_histogram(data = SSAP_pivot %>% filter(Condition == "Sham"),
                 aes(x = Values), bins = 100,
                 color = "blue",
                 alpha = 0.5) +
  geom_histogram(data = SSAP_pivot %>% filter(Condition == "Visual"),
                 aes(x = Values),
                 bins = 100,
                 color = "red",
                 alpha = 0.25) +
  geom_histogram(data = SSAP_pivot %>% filter(Condition == "RCereblm"),
                 aes(x = Values),
                 bins = 100,
                 color = "green",
                 alpha = 0.1)
  
### Example plot difference betweem Pre adn post
ggplot(SSAP_pivot %>% filter(Condition == "Sham" & Post == "PRE")) +
  geom_histogram( aes(x = Values), bins = 100) +
  geom_histogram( data = SSAP_pivot %>% filter(Condition == "Sham" & Post == "POST"),
                  aes(x = Values), bins = 100, 
                  color = "red",
                  alpha = 0.5) 

```
#### Other plots
```{r}

#### Plot per participant pretest
  ggplot(SSAP_pivot %>% 
           filter(Post == "PRE"), aes(x = Condition,
                            y = abs(Values),
                            color = Condition)) +
    geom_point(position = position_jitter()) +
    facet_wrap(~ ID) +
    # geom_smooth(method = "lm") +
    # stat_summary(fun = "mean", colour = "red", size = 2, geom = "point") +
      labs(title = "Absolute values of pitch-to-target difference for each condition per participants",
           subtitle = "Pretest (Melody Imitation)",
         x = "Conditions",
         y = "Absolute values of pitch-to-target difference") +
    theme_bw() +
    theme(text=element_text(size = 10)) +
 scale_x_discrete(labels = c("LCereblm" = "LC", "RCereblm" = "RC", "Sham" = "S", "Visual" = "V")) +
 scale_y_continuous(limits = c(0, 300))

#### Plot per participant posttest
  ggplot(SSAP_pivot %>% 
           filter(Post == "POST"), aes(x = Condition,
                            y = abs(Values),
                            color = Condition)) +
    geom_point(position = position_jitter()) +
    facet_wrap(~ ID) +
    # geom_smooth(method = "lm") +
    # stat_summary(fun = "mean", colour = "red", size = 2, geom = "point") +
      labs(title = "Absolute values of pitch-to-target difference for each condition per participants",
           subtitle = "Posttest (Melody Imitation)",
         x = "Conditions",
         y = "Absolute values of pitch-to-target difference") +
    theme_bw() +
    theme(text=element_text(size = 10)) +
 scale_x_discrete(labels = c("LCereblm" = "LC", "RCereblm" = "RC", "Sham" = "S", "Visual" = "V")) +
 scale_y_continuous(limits = c(0, 300))

#### Pre vs Post
ggplot(data = SSAP_pivot,
       mapping = aes(x = Condition,
                     y = abs(Values),
                     color = Condition)) +
  geom_point(position = position_jitter()) +
  facet_wrap(~ Post) +
  stat_summary(fun = "mean", colour = "red", size = 2, geom = "point") +
    labs(title = "Absolute values of pitch-to-target difference for each condition in melody imitation",
       x = "Conditions",
       y = "Absolute values of pitch-to-target difference") +
  theme_bw() +
  theme(text=element_text(size = 10)) 


#### Dot and whiskers plot Pre vs Post
ggplot(SSAP_pivot) +
  stat_summary(aes(x = Condition,
                   y = abs(Values)),
                   fun.data = "mean_cl_boot", geom = "pointrange") +
      facet_wrap(~ Post) +
    # geom_point(aes(x = Condition,
    #                y = abs(Values), color = Condition),
    #            position = position_jitter(),
    #            alpha = 0.5) +
   labs(title = "Mean absolute value and sd of pitch-to-target difference for each condition 
        \ in melody imitation",
        subtitle = "Pre and Post test",
       x = "Conditions",
       y = "Absolute values of pitch-to-target difference") +
  theme_bw() +
  theme(text=element_text(size = 12))
```



### Transform data for Mplus to try to find auxiliary variables for the model
```{r}
### Create missingness indicator for variable having missing data
SSAP_MPlus <- SSAP_pivot %>% mutate(R_Values = if_else(is.na(Values),1,0 ))

### Change categorical variables to number for Mplus.
SSAP_MPlus <- SSAP_MPlus %>% mutate(Gender = if_else(str_detect(Gender, pattern = "M"), 1, 0),
                      Post = if_else(str_detect(Post, pattern = "POST"), 1, 0),
                      Condition = case_when(str_detect(Condition, pattern = "Sham") ~ 0,
                                            str_detect(Condition, pattern = "RCereblm") ~ 1,
                                            str_detect(Condition, pattern = "LCereblm") ~ 2,
                                            TRUE ~ 3))

### Transform NA data to -99 for Mplus format
SSAP_MPlus <- SSAP_MPlus %>% 
  mutate(across(everything(),
                ~replace_na(.x, -99))) %>% 
  select(!ID & !Last_practiced)

#### Save as csv file for Mplus
write_csv(SSAP_MPlus, file = "SSAP_MPlus.csv",col_names = F, append = F)
```

### Start multiple imputation process

* Used [this tutorial](https://bookdown.org/marklhc/notes_bookdown/missing-data.html) for missing data
```{r}
### Create missing indicator value for missing data in Values
SSAP_pivot_R <- transform(SSAP_pivot,
                          Values_R = factor(as.numeric(!is.na(Values)),
                                            labels = c("missing",
                                                       "observed")))
# ### Plot 
#  ggplot(SSAP_pivot_R) +
#    geom_col(aes(x = Values_R, y = Condition, file = Condition))


### Look at the default setting
ImpSSAP <- mice(SSAP_pivot, m = 1, print = F)
meth <- ImpSSAP$meth # Check Imputation method and save it if happy with it (Default is pmm: Predictive mean matching)
pred <- ImpSSAP$pred # Check predictor matrix and save it 


### Set variables that have no predictive power (e.g. ID) to zero in predictor matrix
pred[ ,"ID"] <- 0 # Set column ID to zero
pred[, "Gender"] <-  0 # Set column Gender to zero

### Step 1: Imputation 
ImpSSAP_new <- mice(SSAP_pivot, m = 20, method = meth, pred = pred, print = F, maxit = 50, seed = 29002)
 
### Print trace plot to check for convergence

png("myplot.png")
myplot <- plot(ImpSSAP_new)
print(myplot)
dev.off()

## Plot real data against imputed data
png("myplot2.png")
myplot2 <- stripplot(ImpSSAP_new, Values ~.imp, pch=20, cex=2)
print(myplot2)
dev.off()

### Base version with other variables
stripplot(ImpSSAP_new)
### Other way to plot if have multiple variables with missing data
# xyplot(tempData,Ozone ~ Wind+Temp+Solar.R,pch=18,cex=1)

### Plot density plot
png("ppcheck2.png")
ppcheck2 <- densityplot(ImpSSAP_new)
print(ppcheck2)
dev.off()
### Check imputed data set
complete(ImpSSAP_new, "long") %>% head()

#### Get prior
get_prior(Values ~ 1 + Condition + Post + Condition:Post + (1 + Condition + Post|ID) + (1 | Interval), # might need to specify combine = TRUE so that the brmsfit_multiple object, which inherits from class brmsfit and behaves essentially the same as brmsfit object.
                            data = ImpSSAP_new)

### Fit the bayesion model
m1_SSAP_imp <- brm_multiple(Values ~ 1 + Condition + Post + Condition:Post + (1 + Condition + Post|ID) + (1 | Interval), # might need to specify combine = TRUE so that the brmsfit_multiple object, which inherits from class brmsfit and behaves essentially the same as brmsfit object.
                            data = ImpSSAP_new,
                            prior = c(prior(normal(0, 20), class = "Intercept"),
                                 # set for all "b" coefficients
                                 prior(normal(0, 10), class = "b"),
                                 prior(normal(0, 50), class = "sigma")),
                                 # prior(student_t(4, 0, 1), class = "sigma")),
                       seed = 29002,
                       chains = 4L,
                       cores = 4,
                       file = "imp_mel_SSAP_data"
)


### Check if each model converged: Look at rhat for all models
m1_SSAP_imp$rhats

### Pool results together
pooled_SSAP <- combine_models(m1_SSAP_imp)

### Check poolded model
summary(pooled_SSAP)

### Compare results with model that remove all missing values
m1_SSAP_reg <- brm(Values ~ Condition * Post + (1 + Condition | ID),
                  data = SSAP_pivot,
                  prior = c(prior(normal(0, 1), class = "Intercept"),
                                 # set for all "b" coefficients
                                 prior(normal(0, 1), class = "b"),
                                 prior(student_t(4, 0, 1), class = "sigma")))

m1_SSAP_reg <- brm(Values ~ 1 + Condition + Post + Condition:Post + (1 + Condition + Post|ID) + (1 | Interval), # might need to specify combine = TRUE so that the brmsfit_multiple object, which inherits from class brmsfit and behaves essentially the same as brmsfit object.
                            data = SSAP_pivot,
                            prior = c(prior(normal(0, 20), class = "Intercept"),
                                 # set for all "b" coefficients
                                 prior(normal(0, 10), class = "b"),
                                 prior(normal(0, 50), class = "sigma")),
                                 # prior(student_t(4, 0, 1), class = "sigma")),
                       seed = 29002,
                       chains = 4L,
                       cores = 4,
                    file = "SSAP_data_reg")

summary(m1_SSAP_reg)

```


```{r}
## Summary table for MI
### Create data frame with needed values 
dfmi <- data.frame(Pop_Level_Effects = c("Intercept", "Condition1", "Condition2", "Condition3", "Post", 
                                         "Condition1:Post", "Condition2:Post", "Condition3:Post"),
           estimate = c(16.64, -8.61, -3.30, -1.53, 2.62, 0.56, -6.43, -3.96),
           est_error = c(9.57, 5.39, 5.19, 5.25, 4.50, 6.09, 5.78, 6.20),
           l_95_CI = c(-3.20, -18.93, -13.48, -11.84, -6.40, -11.65, -17.63, -15.90),
           u_95_CI = c(34.62, 2.25, 6.94, 8.81, 11.32, 12.33, 4.95, 8.38),
           PSRF = c(1.00, 1.04, 1.01, 1.08, 1.01, 1.05, 1.00, 1.08))
           
          

### Create table
dfmi %>% 
  kableExtra::kbl(caption = "Summary table Bayesian analysis after pooled results") %>% 
  kableExtra::kable_classic(full_width = F, html_font = "Time") %>% 
  kableExtra::save_kable("dfsaved.png") 
```


```{r}
posterior_imp <- as.matrix(m1_SSAP_imp)

mcmc_areas(posterior_imp,
           pars = c(
             "b_Intercept", "b_Condition1","b_Condition2", "b_Condition3", "b_Post1"), 
                    # "b_Condition1:Post1","b_Condition2:Post1","b_Condition3:Post1"),
           point_est = c("mean"),
           prob = 0.95) +
  labs(title = "Posterior distributions with means and 95% credibility intervals for melody imitation \
  imputed dataset",
         x = "Deviation (in cents)",
         y = "Variables") +
   theme(text=element_text(size = 10)) 
ggsave("mcmc_post_nel.png")
```


#### Build tables from results
```{r}
### Summary table for MI
### Create data frame with needed values 
df <- data.frame(Pop_Level_Efct = c("Intercept", "ConditionRCereblm", "ConditionSham", "ConditionVisual               ", "PostPRE", "ConditionRCereblm:PostPRE", "ConditionSham:PostPRE", "ConditionVisual:PostPRE"),
           est = c(0.22, -0.06, 0.21, 0.07, 0.03, 0.15, -0.00, 0.19),
           est_error = c(1.19, 0.98, 1.00, 0.99, 0.97, 0.97, 0.98, 1.00),
           l_95_CI = c(-2.10, -1.96, -1.71, -1.86, -1.86, -1.73, -1.95, -1.75),
           u_95_CI = c(2.52, 1.88, 2.15, 2.01, 1.95, 2.07, 1.92, 2.18),
           PSRF = c(1.00, 1.00, 1.00, 1.00, 1.01, 1.00, 1.00, 1.01))
           
          

### Create table
df %>% 
  kableExtra::kbl(caption = "Summary table Bayesian analysis after pooled results") %>% 
  kableExtra::kable_classic(full_width = F, html_font = "Cambria") %>% 
  kableExtra::save_kable("dfsaved.png") 



### Summary table for complete data after removing missing
df2 <- data.frame(Pop_Level_Efct = c("Intercept", "ConditionRCereblm", "ConditionSham", "ConditionVisual               ", "PostPRE", "ConditionRCereblm:PostPRE", "ConditionSham:PostPRE", "ConditionVisual:PostPRE"),
           est = c(0.20, 0.04, 0.26, 0.08, -0.04, 0.14, -0.00, 0.12),
           est_error = c(1.21, 0.97, 1.00, 1.00, 0.96, 0.98, 0.99, 1.02),
           l_95_CI = c(-2.20, -1.89, -1.67, -1.80, -1.88, -1.78, -1.94, -1.87),
           u_95_CI = c(2.53, 1.97, 2.21, 2.05, 1.83, 2.00, 1.87, 2.10),
           PSRF = c(1.00, 1.00, 1.01, 1.00, 1.00, 1.00, 1.00, 1.00))


### Create table
df2 %>% 
  kableExtra::kbl(caption = "Summary table Bayesian analysis after listwise deletion") %>% 
  kableExtra::kable_classic(full_width = F, html_font = "Cambria") %>% 
  kableExtra::save_kable("df2saved.png")

summary(m1_SSAP_reg)
```


```{r}
sessionInfo()
```

